### 2019-11-25 RFCs/HDF5/VOL/connector_author_guide/stacking.tex
Most API changes in. Still need to move "helper" text over.
Former-commit-id: c236ad85c2fcb331051a2de3ffa110e6bfaeb004

\section{Interchanging and Stacking VOL Plugins}
\label{sec:stack}

@@

\subsection{Stacking Plugins on Top of Each Other}
It would also be possible to stack VOL plugins on top of each
other. This notion is similar to the idea of the split VFD, where
underneath the split VFD itself, two file drivers would be used, one
for the file storing the metadata and another for raw data. Some
stackings make sense and others would be erroneous. For example,
stacking the native HDF5 plugin on top of a non-HDF5 backend plugin
does not make sense and is erroneous. Figure~\ref{fig:stack} shows a
stacking of a remote plugin, where data is distributed remotely, on
top of the native h5 plugin, where servers that store the data at
remote locations use the h5 file format.

@@

\subsection{Mirroring Plugins}
Another useful design option is to allow a mirroring plugin, where the
HDF5 API calls are forwarded through a mirror plugin to two or more
VOL plugins. This is an extension to the stacking
feature. Figure~\ref{fig:mirror} shows an example of a VOL mirror that
maps HDF5 API calls to an h5 backend plugin and an XML backend plugin.

@@

\subsection{Implementing Stacked \& Mirrored Plugins}
A new set of API calls that map directly to the VOL callbacks have
been added to the HDF5 library to make stacking and mirroring easy for
plugin developers. Those APIs are listed at then end of Section~\ref{sec:api}. They are similar to the public VFD (H5FD) routines that call the VFD callbacks directly.

>>>>

\section{Interchanging and Stacking VOL Connectors}
\label{sec:stack}

@@

\subsection{Stacking Connectors on Top of Each Other}
It would also be possible to stack VOL plugins on top of each
other. This notion is similar to the idea of the split VFD, where
underneath the split VFD itself, two file drivers would be used, one
for the file storing the metadata and another for raw data. Some
stackings make sense and others would be erroneous. For example,
stacking the native HDF5 plugin on top of a non-HDF5 backend plugin
does not make sense and is erroneous. Figure~\ref{fig:stack} shows a
stacking of a remote plugin, where data is distributed remotely, on
top of the native h5 plugin, where servers that store the data at
remote locations use the h5 file format.

@@

\subsection{Mirroring Connectors}
Another useful design option is to allow a mirroring plugin, where the
HDF5 API calls are forwarded through a mirror plugin to two or more
VOL plugins. This is an extension to the stacking
feature. Figure~\ref{fig:mirror} shows an example of a VOL mirror that
maps HDF5 API calls to an h5 backend plugin and an XML backend plugin.

@@

\subsection{Implementing Stacked \& Mirrored Connectors}
A new set of API calls that map directly to the VOL callbacks have
been added to the HDF5 library to make stacking and mirroring easy for
plugin developers. Those APIs are listed at then end of Section~\ref{sec:api}. They are similar to the public VFD (H5FD) routines that call the VFD callbacks directly.

### 2019-11-25 RFCs/HDF5/VOL/connector_author_guide/writing_new.tex
Most API changes in. Still need to move "helper" text over.
Former-commit-id: c236ad85c2fcb331051a2de3ffa110e6bfaeb004

\section{Creating and Using an External Plugin}
External plugins are developed outside of the HDF5 library and do not
use any internal HDF5 private functions. They do not require to be
shipped with the HDF5 library, but can just link to it from userspace
just like an HDF5 application. If an external VOL plugin is installed on the system as a shared library or a DLL, the library can search for it and load it dynamically to be used by an application.

\subsection{Creating an External Plugin}
Developing an external plugin is similar to developing an internal plugin. Refer to Section~\ref{sec:vol} for details on plugin creation. The important thing to keep in mind is that external plugins cannot use internal HDF5 features. The ``value'' field for an external plugin should be a positive integer greater than 128. The name should be unique across all plugins registered with the library.

@@

\subsection{Using an External Plugin}
Unlike internal plugins, external plugins cannot create an API routine for applications to use to set the VOL plugin in the file access property list. External plugin developers can, however, provide a wrapper routine that registers the plugin with {\tt H5VLregister()} instead of having the application call that directly with the pointer to the plugin class structure. Both ways are possible if the external plugin used is in the application's userspace. External plugins provided by third party vendors that are not in the application's userspace but are installed on the system with a shared library or DLL can be registered by the application using {\tt H5VLregister\_name()} provided that the application knows the plugin's name in advance.

@@

% /* Function prototypes */
% H5_DLL htri_t H5VLis_registered(hid_t id);
% H5_DLL ssize_t H5VLget_plugin_name(hid_t id, char *name/*out*/, size_t size);

% H5_DLL void *H5Pget_vol_info(hid_t plist_id);
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\section{Creating a New Connector}
External plugins are developed outside of the HDF5 library and do not
use any internal HDF5 private functions. They do not require to be
shipped with the HDF5 library, but can just link to it from userspace
just like an HDF5 application. If an external VOL plugin is installed on the system as a shared library or a DLL, the library can search for it and load it dynamically to be used by an application.

\subsection{Overview}
Developing an external plugin is similar to developing an internal plugin. Refer to Section~\ref{sec:vol} for details on plugin creation. The important thing to keep in mind is that external plugins cannot use internal HDF5 features. The ``value'' field for an external plugin should be a positive integer greater than 128. The name should be unique across all plugins registered with the library.

@@

\subsection{Using The New Connector}
Unlike internal plugins, external plugins cannot create an API routine for applications to use to set the VOL plugin in the file access property list. External plugin developers can, however, provide a wrapper routine that registers the plugin with {\tt H5VLregister()} instead of having the application call that directly with the pointer to the plugin class structure. Both ways are possible if the external plugin used is in the application's userspace. External plugins provided by third party vendors that are not in the application's userspace but are installed on the system with a shared library or DLL can be registered by the application using {\tt H5VLregister\_name()} provided that the application knows the plugin's name in advance.

@@

\subsection{Testing}

\subsection{Internal Connectors}

Internal plugins are developed internally with the HDF5 library and
are required to ship with the entire library to be used. Typically
those plugins need to use internal features and functions of the HDF5
library that are not available publicly from the user application.

\subsubsection{Implementing an Internal Connector}
The first step to implement an internal plugin is to implement all the callbacks defined in the VOL class as described in Section~\ref{sec:vol}. After implementing the VOL class, the next step would be to allow users to select
this plugin to be used. This is done by creating a new API routine to
set the plugin on the file access property list. For example, if we
create an internal plugin called ``dummy'' that needs an MPI
communicator and info object as information from the user, that
routine signature should look like:

\begin{lstlisting}
herr_t H5Pset_fapl_dummy(hid_t fapl_id, MPI_Comm comm, MPI_Info info);
\end{lstlisting}

The implementation for the above routine should use the internal
function:
\begin{lstlisting}
herr_t H5P_set_vol(H5P_genplist_t *plist, H5VL_class_t *vol_cls, const
void *vol_info);
\end{lstlisting}
that will set the file access using that {\tt fapl\_id} to go through
the ``dummy'' plugin. It will also call the copy callback of the
``dummy'' plugin on the info object ({\tt comm and info}).

A sample implementation for the {\tt H5Pset\_fapl\_dummy()} would look like this:

\begin{lstlisting}
/* DUMMY-specific file access properties */
typedef struct H5VL_dummy_fapl_t {
    MPI_Comm            comm;     /* communicator       */
    MPI_Info            info;     /* MPI information   */
} H5VL_dummy_fapl_t;

herr_t
H5Pset_fapl_dummy(hid_t fapl_id, MPI_Comm comm, MPI_Info info)
{
    H5VL_dummy_fapl_t fa;
    H5P_genplist_t  *plist;      /* Property list pointer */
    herr_t          ret_value;

    FUNC_ENTER_API(FAIL)

    if(fapl_id == H5P_DEFAULT)
        HGOTO_ERROR(H5E_PLIST, H5E_BADVALUE, FAIL, "can't set values in default property list")

    if(NULL == (plist = H5P_object_verify(fapl_id, H5P_FILE_ACCESS)))
        HGOTO_ERROR(H5E_ARGS, H5E_BADTYPE, FAIL, "not a file access property list")

    if(MPI_COMM_NULL == comm)
	HGOTO_ERROR(H5E_PLIST, H5E_BADTYPE, FAIL, "not a valid communicator")

    /* Initialize driver specific properties */
    fa.comm = comm;
    fa.info = info;

    ret_value = H5P_set_vol(plist, &H5VL_dummy_g, &fa);

done:
    FUNC_LEAVE_API(ret_value)
} /* end H5Pset_fapl_dummy() */
\end{lstlisting}

At this point, the internal plugin is ready to be used. For more
information on how to implement an internal plugin, the native plugin
for the HDF5 library is a comprehensive plugin that implements all
features of the library and can be used as a guide.

\subsubsection{Using an Internal Connector}
An internal plugin is registered automatically with the HDF5 library when the library gets initialized. Users should not attempt to register an internal plugin. The internal registration operation creates a global identifier for each internal plugin. Applications can query the library for that identifier using the plugin name with {\tt H5VLget\_plugin\_id()}. This identifier is used to initialize the plugin if it requires an initialization phase. A third party library could use that ID to create a plugin that stacks or mirrors on top of an internal plugin using that identifier (refer to Section~\ref{sec:stack} for more details on stacking and mirroring).

Some plugins could require an initialization phase before using the plugin with some input parameters from users. For example, a remote plugin would need to setup a connection to a remote machine with a specific hostname that the user knows and the HDF5 library does not know about. This process is done by creating a VOL initialization property list and setting a public property that should be defined by the internal plugin to the hostname of the remote machine. The user then calls the initialize routine ({\tt H5VLintialize()}) with the plugin\_id and the property list to initialize the plugin. Some plugins could create a public API routine that wraps the property list creation, property settings, and plugin initialization into one routine that the user calls to simplify this process for the users. The native HDF5 plugin does not require an initialization phase.

Now that the plugin is initialized, the application can create or open an HDF5 container using the internal plugin by creating a file access property list and calling the corresponding API routine that the internal plugin should define for setting the VOL plugin property and container parameters on the file access property list. That FAPL can be passed to the {\tt H5Fcreate()} or {\tt H5Fopen{}} calls to access the container through the VOL plugin. Subsequently, all other HDF5 calls on the container and objects created or opened in that container will automatically go to the selected VOL plugin.

After closing all objects and containers on that plugin, the application should terminate access to the selected plugin if it requires a termination phase by calling {\tt H5VLterminate()} with the required termination properties (similar to the initialization phase). The native HDF5 plugin does not require a termination phase.

The internal plugin is unregistered automatically by the HDF5 library on library termination, and users should not attempt to unregister it themselves.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:
